A Standard for the Scholarly Citation of Archaeological Data

### 2017-09-07 analysis/paper/paper.Rmd
add a few citations

```{r date-and-sharing, fig.cap="There is no obvious trend in data sharing becavious and date of publication", eval = TRUE}
library(viridis)
years <- seq(min(email_survey$'publication year', na.rm = TRUE),
    max(email_survey$'publication year', na.rm = TRUE), 1)
email_survey %>% 
  filter(!is.na( `email sent date`)) %>% 
  group_by(`publication year`, `content recieved date`) %>% 
  tally() %>% 
  filter(!is.na(`publication year`)) %>% 
  mutate(`content recieved date` = if_else(!is.na(`content recieved date`), 
  "yes", "no")) %>% 
  ggplot(aes(`publication year`, n)) +
    geom_col(aes(fill = `content recieved date`),
             position = position_dodge(preserve = 'single')) +
    scale_fill_viridis(discrete = TRUE,
                       guide = guide_legend(title = "Data shared?")) +
    scale_x_continuous(breaks = years,
                       labels = years) +
    theme_bw()

@@

From this pilot study of repository data we see that policy mandates in the UK and the Netherlands have been effective in populating repositories with archaeological data. Nearly all records are made available under conditions that permit others to use the data while requiring that the original author of the data receive credit in reuse. In most cases this is because the repository hosting the data requires the author to choose one of these permissive licenses as a condition of making the deposit. However, although many items are _able_ to be reused because of the licenses, then ease of reuse of these items is low. Only a small fraction of the items deposited are machine-readable structured data files. The majority of items are PDFs, which require extensive manual handling to extract tablular data for convenient reuse and combination with other datasets.

@@

There are many guidelines and conventions from other disciplines that are relevant to sharing archaeological data. Table XXX summarises some of these, and recommendations specifically written for archaeologists can be found in XXX [archaeology data service manuals]. We recognise that many of these guidelines may create new technical and practical challenges to archaeologists who are not accustomed to sharing data, and many not want to dedicate resources to sharing data. However, we should not let the perfect be the enemy of the good. In our own work we try to follow these guidelines with no dedicated resources for data sharing, and with an incomplete knowledge of some of the technical details of repository operations. While we recognise that our data sharing pracitses fall short of all the guidelines, we still see value in sharing data imperfectly, rather than not at all.

@@

<!-- 
In the pilot studies reported above we see three common approaches to sharing data: sending files privately by email, following a personal request; including data as supplementary files submitted for publication with a journal article; and sharing data by depositing files in a trustworthy, public, DOI-issuing repository. The third method, via a public repository, is by far the most widely recommended method both specifically by archaeologists (Kintigh 2006; Kansa et al. 2011; Kansa and Kansa 2013), and in other sciences (cf. Cranston et al. 2014; Penev et al. 2017; Mounce 2014; Thessen & Patterson 2011; Whitlock 2011).		
 -		
 -		
 -Several guides with detailed information relevant to preparing archaeological data for sharing on a repository. 		
 -		
 -- http://guides.archaeologydataservice.ac.uk/		
 -- http://archaeologydataservice.ac.uk/arches/Wiki.jsp?page=ARCHES%20Introduction		
 -		
 -- http://www.icpsr.umich.edu/files/ICPSR/access/dataprep.pdf		
 -- https://www.dataone.org/sites/all/documents/DataONE_BP_Primer_020212.pdf		
 -- http://www.data-archive.ac.uk/media/2894/managingsharing.pdf
 -->

@@

@@

@@

@@

@@

>>>>

```{r date-and-sharing, fig.cap="There is no obvious trend in data sharing becavious and date of publication", eval = TRUE}
library(viridis)
years <- seq(min(email_survey$'publication year', na.rm = TRUE),
    max(email_survey$'publication year', na.rm = TRUE), 1)
email_survey %>% 
  filter(!is.na( `email sent date`)) %>% 
  group_by(`publication year`, `content recieved date`) %>% 
  tally() %>% 
  filter(!is.na(`publication year`)) %>% 
  mutate(`content recieved date` = if_else(!is.na(`content recieved date`), 
  "yes", "no")) %>% 
  ggplot(aes(`publication year`, n)) +
    geom_col(aes(fill = `content recieved date`),
             # need devtools::install_github("tidyverse/ggplot2") for this
             position = position_dodge(preserve = 'single')) +
    scale_fill_viridis(discrete = TRUE,
                       guide = guide_legend(title = "Data shared?")) +
    scale_x_continuous(breaks = years,
                       labels = years) +
    theme_bw()

@@

From this pilot study of repository data we see that policy mandates in the UK and the Netherlands have been effective in populating repositories with archaeological data. With the 2012 spike in records deposited in tDAR we also see that archaeologists are sensitive to the price of sharing data. Nearly all records are made available under conditions that permit others to use the data while requiring that the original author of the data receive credit in reuse. In most cases this is because the repository hosting the data requires the author to choose one of these permissive licenses as a condition of making the deposit. However, although many items are _able_ to be reused because of the licenses, then ease of reuse of these items is low. Only a small fraction of the items deposited are machine-readable structured data files. The majority of items are PDFs, which require extensive manual handling to extract tablular data for convenient reuse and combination with other datasets.

@@

In the pilot studies reported above we see three common approaches to sharing data: sending files privately by email, following a personal request; including data as supplementary files submitted for publication with a journal article; and sharing data by depositing files in a trustworthy, public, DOI-issuing repository. The third method, via a public repository, is by far the most widely recommended method both specifically by archaeologists (Kintigh 2006; Kansa et al. 2011; Kansa and Kansa 2013), and in other sciences (cf. Cranston et al. 2014; Penev et al. 2017; Mounce 2014; Thessen & Patterson 2011; Whitlock 2011).	There are many guidelines and conventions from other disciplines that are relevant to sharing archaeological data (ICPSR 2012; Strasser et al. 2012). Table XXX summarises some of these, and recommendations specifically written for archaeologists can be found in UK Archaeology Data Service (n.d.), ARCHES project. (n.d.) [archaeology data service manuals]. Many of these guidelines may create new technical and practical challenges to archaeologists who are not accustomed to sharing data, and many not want to dedicate resources to sharing data. However, we should not let the perfect be the enemy of the good. In our own work we try to follow these guidelines with no dedicated resources for data sharing, and with an incomplete knowledge of some of the technical details of repository operations. While we recognise that our data sharing pracitses fall short of all the guidelines, we still see value in sharing data imperfectly, rather than not at all.

@@

@@

 # Ethical limitations to data sharing

@@

ARCHES project. (n.d.). The Standard and Guide to Best Practice in Archaeological Archiving in Europe. Retrieved September 7, 2017, from http://archaeologydataservice.ac.uk/arches/Wiki.jsp?page=ARCHES%20Introduction

@@

Inter-university Consortium for Political and Social Research (ICPSR). (2012). Guide to Social Science Data Preparation and Archiving: Best Practice Throughout the Data Life Cycle (5th ed.). Ann Arbor, MI

@@

Strasser, C., Cook, R., Michener, W., & Budden, A. (2012). Primer on Data Management: What you always wanted to know. EScholarship. https://doi.org/10.5060/D2251G48

@@

UK Archaeology Data Service. (n.d.). Guides to Good Practice: Main. Retrieved September 7, 2017, from http://guides.archaeologydataservice.ac.uk/
